---
title: "Analysis in a Box"
subtitle: "A Practical Guide to Using RStudio with Docker"
author:
  - name: Ismael Gutiérrez-Cordero
    affiliation: Universidad de Málaga
  - name: Bernhard Angele
    affiliation: Universidad Antonio de Nebrija
date: today
format:
  revealjs:
    theme: default
    slide-number: true
    toc: true
    toc-depth: 1
    toc-title: "Overview"
    logo: ""
    footer: "Analysis in a Box | RStudio + Docker"
    code-fold: false
    code-tools: true
    highlight-style: github
    transition: slide
    incremental: false
    chalkboard: true
    css: styles.css
bibliography: references.bib
csl: apa.csl
---

# The Problem

## Modern Analyses Are Complex

::: {.incremental}
- Behavioral and neuroscientific data analysis has become **increasingly complex**
- ANOVAs are no longer sufficient — researchers now use:
  - (Generalized) Linear Mixed Models (GLMMs) via **lme4** [@bates2015fitting]
  - Bayesian multilevel models via **brms** [@burkner2017brms]
- *brms* requires a large number of dependencies:
  - Stan [@stan2025] for MCMC sampling
  - CmdStanR [@gabry2025] (recommended) or RStan [@rstan2025]
  - A modern C++ compiler (RTools on Windows, XCode on Mac)
:::

## Installation Challenges

:::: {.columns}

::: {.column width="50%"}
**New Users**

- Overwhelmed by complex installation chains
- Multiple tools to install before writing a single line of R
- Different instructions per operating system
:::

::: {.column width="50%"}
**Experienced Users**

- Significant time spent re-installing and updating packages
- Package updates can break existing code
- Replicating analyses from even a few years ago can fail
:::

::::

## The Reproducibility Gap

> *"Computational results cannot be comprehensively evaluated without knowing the code used to perform the computations."*
> — @ince2012case

::: {.incremental}
- Package versions change constantly, breaking analysis code
- Solutions like **renv** [@renv] require significant effort from both authors and readers
- We need a simpler, more robust approach...
:::

# Docker to the Rescue

## What Is Docker?

::: {.callout-note}
**Docker** is an open-source platform that allows users to run applications in isolated environments called **containers**.
:::

Key concepts:

| Term | Meaning |
|------|---------|
| **Image** | A pre-built software package with all code and configuration |
| **Container** | A running instance of an image — isolated from the host |
| **Docker Hub** | Public repository for sharing Docker images |

## Why Docker for R Analyses?

::: {.incremental}
- **Consistency**: The same image runs identically on any machine
- **Reproducibility**: Freeze your exact software environment as an image
- **Ease of setup**: One command to get a fully working R + RStudio + brms environment
- **Shareability**: Push your image to Docker Hub; anyone can pull and run it
- **Versioning**: Images can build on each other — add packages to an existing base
:::

## Available Base Images

:::: {.columns}

::: {.column width="50%"}
**🐳 Rocker Project** [@rocker2017]

- <https://rocker-project.org>
- R + RStudio + tidyverse/verse
- Versioned by R version
- Best for **Intel/AMD64** (Windows/Linux)
- `rocker/verse:4.5.3`
:::

::: {.column width="50%"}
**🐳 b-data Images**

- <https://gitlab.com/b-data>
- R + RStudio + tidyverse/verse  
- Supports **both amd64 AND arm64**
- Best for **Apple M-series** and ARM Windows
- Recommended for maximum performance on ARM
:::

::::

## Our "Analysis in a Box" Image

We built a custom image on top of Rocker verse that includes:

- base R + RStudio Server
- tidyverse and data processing packages
- **brms** with CmdStanR (pre-compiled!)
- lme4, buildmer, emmeans
- Quarto + apaquarto template
- LaTeX / TinyTeX for PDF rendering

**Available on Docker Hub:**  
<https://hub.docker.com/r/bangele1/analysis-in-a-box-rocker>

# Getting Started

## Prerequisites

1. **Install Docker Desktop**

   Follow the official guide: <https://docs.docker.com/get-started/introduction/get-docker-desktop/>

2. **Verify Docker is working:**

   ```bash
   docker run -d -p 8081:80 docker/welcome-to-docker
   ```

   Then open <http://localhost:8081> in your browser.

3. **Create a working folder** (where your files will be saved):

   ::: {.panel-tabset}
   ### Windows CMD
   ```cmd
   mkdir %USERPROFILE%\Documents\analysis_box
   cd %USERPROFILE%\Documents\analysis_box
   ```

   ### macOS / Linux
   ```bash
   mkdir -p ~/Documents/analysis_box
   cd ~/Documents/analysis_box
   ```
   :::

## Running the Container

Navigate to your working folder and run:

::: {.panel-tabset}

### Windows CMD

```cmd
docker run -e PASSWORD=yourpassword -p 8787:8787 ^
  -v "%cd%":/home/rstudio ^
  bangele1/analysis-in-a-box-rocker:latest
```

### macOS / Linux

```bash
docker run -e PASSWORD=yourpassword -p 8787:8787 \
  -v $(pwd):/home/rstudio \
  bangele1/analysis-in-a-box-rocker:latest
```

:::

. . .

What do these options mean?

- **`-e PASSWORD=yourpassword`** — sets the RStudio Server login password
- **`-p 8787:8787`** — exposes port 8787 to your browser
- **`-v $(pwd):/home/rstudio`** — mounts your local folder inside the container

## Accessing RStudio

1. Open your browser and navigate to **<http://127.0.0.1:8787>**

2. Log in with:
   - **Username:** `rstudio`
   - **Password:** *(whatever you set after `PASSWORD=`)*

3. You now have a **fully working RStudio environment** with brms, CmdStanR, and all dependencies pre-installed!

::: {.callout-tip}
The CLI/terminal window must remain open for the container to stay running.
:::

## Downloading the Example Files

Inside RStudio, open the **Terminal** tab and run:

```bash
git clone https://github.com/bernhardangele/docker_presentation.git
```

This downloads the example data and analysis scripts.  
Open `docker_presentation.RProj` to load the project in RStudio.

## Using Docker Compose

For more complex setups, use **Docker Compose** instead of a long `docker run` command.

Create a file called `docker-compose.yml`:

```yaml
services:
  analysis-in-a-box:
    image: bangele1/analysis-in-a-box-rocker:latest
    environment:
      - PASSWORD=yourpassword
    ports:
      - "8787:8787"
    volumes:
      - .:/home/rstudio
```

Then run:

```bash
docker compose up -d     # Start container in background
docker compose down      # Stop and remove container
```

# Analyses with Docker

## Example Dataset

A lexical decision experiment with a flanker manipulation:

:::: {.columns}

::: {.column width="50%"}
**Design**

- **Stimulus type**: words vs. pseudowords
- **Gender**: feminine vs. masculine targets
- **Condition**: concordant (AGR) vs. discordant (DIS) flanker articles
- Participants and items crossed (fully counterbalanced)
:::

::: {.column width="50%"}
**Outcome measures**

- Response times (RT) in seconds
- Accuracy (correct lexical decisions)

**Key question:** Do gender-congruent flankers facilitate lexical decisions?
:::

::::

## Example Trial Display

Each trial shows a target word (or pseudoword) flanked by articles.

:::: {.columns}

::: {.column width="50%"}

**Concordant (AGR)**

Target word *mesa* (feminine) — flanker *la* (feminine)

<div style="width:340px; height:180px; border:3px solid #000; border-radius:8px; background:#fff; display:flex; align-items:center; justify-content:center; margin:10px auto;">
  <span style="color:#000; font-size:28px; font-family:monospace; letter-spacing:8px;">la&nbsp;&nbsp;mesa&nbsp;&nbsp;la</span>
</div>

Press **L** for feminine
:::

::: {.column width="50%"}

**Discordant (DIS)**

Target word *mesa* (feminine) — flanker *el* (masculine)

<div style="width:340px; height:180px; border:3px solid #000; border-radius:8px; background:#fff; display:flex; align-items:center; justify-content:center; margin:10px auto;">
  <span style="color:#000; font-size:28px; font-family:monospace; letter-spacing:8px;">el&nbsp;&nbsp;mesa&nbsp;&nbsp;el</span>
</div>

Press **A** for masculine
:::

::::

## Mixed-Effects Models: Why?

Standard ANOVA treats participants (or items) as the only random source of variability.  
Mixed models include **random effects** for both:

:::: {.columns}

::: {.column width="33%"}
<div style="font-size:0.8em;">

**Random intercepts only**

Each participant has a different baseline RT, but the *effect* of condition is the same for everyone.

```r
rt ~ Condition * Gender + 
  (1 | PROLIFIC_PID) + 
  (1 | Target)
```
</div>
:::

::: {.column width="33%"}
<div style="font-size:0.8em;">

**Random intercepts + slopes**

Each participant also has a different *effect* of condition.

```r
rt ~ Condition * Gender + 
  (Condition * Gender | PROLIFIC_PID) +
  (Condition | Target)
```
</div>
:::

::: {.column width="33%"}
<div style="font-size:0.8em;">

**Why it matters**

- Avoids pseudoreplication
- By-item random effects: items differ in difficulty
- Converges to correct Type I error rates
</div>
:::

::::

## Frequentist Analysis: lme4 + buildmer

```{r}
#| cache: true
#| 
library(lme4)
library(buildmer)
library(tidyverse)

# Read all PsychoJS trial data
exp2_all_csv_content <- fs::dir_ls(path = "../data", glob = "*.csv") %>%
  map_dfr(read_csv, .id = "source", col_type = cols(
    .default = col_character(), rt = col_double(), corr = col_integer(), TrialID = col_integer()))

exp2 <- exp2_all_csv_content |>
  group_by(PROLIFIC_PID) |>
  mutate(counterbalancing_condition = counterbalancing_condition[1]) |>
  ungroup() |>
  filter(!is.na(TrialID) & TrialID < 2000) |>
  mutate(Target = Prime) |>
  filter(rt > .25 & rt < 1.8 & corr == 1 & StimulusType == "WORD")

exp2$Condition <- factor(exp2$Condition, levels = c("AGR", "DIS"))
exp2$Gender <- factor(exp2$Gender, levels = c("FEM", "MAS"))

# Fit model using buildmer
model_rt <- buildmer(
  rt ~ Condition * Gender +
    (Condition * Gender | PROLIFIC_PID) +
    (Condition | Target),
  data = exp2,
  buildmerControl = buildmerControl(direction = "order")
)

summary(model_rt)
```

::: {.callout-note}
**buildmer** [@voeten2025buildmer] automatically reduces the random effects structure to find the maximal converging model — saving considerable trial and error.
:::

## Bayesian Analysis: brms (pre-computed)

```{r}
#| cache: true
library(brms)
library(qs2)
library(gt)
library(tidyverse)

# Load pre-computed models (run in analysis_exp2.R)
blmm_exp2_rt_dist <- qs_read("../analysis/blmm_exp2_rt_dist.qs")
blmm_acc_exp2 <- qs_read("../analysis/blmm_acc_exp2.qs")
```

<div style="font-size:0.55em;">

Models were fit using an **exgaussian** distribution for RT:

```r
# RT model (words only, correct responses, 0.25s < RT < 1.8s)
blmm_exp2_rt_dist <- brm(
  data = exp2 |> filter(rt > .25 & rt < 1.8 & corr == 1 & StimulusType == "WORD"),
  formula = bf(
    rt ~ Condition * Gender + (Condition * Gender | PROLIFIC_PID) + (Condition | Target),
    beta ~ Condition * Gender + (Condition * Gender | PROLIFIC_PID) + (Condition | Target)
  ),
  family = exgaussian(), chains = 4, iter = 5000, backend = "cmdstanr"
)

# Accuracy model (Bernoulli logistic regression)
blmm_acc_exp2 <- brm(
  data = exp2 |> filter(rt > .25 & rt < 1.8 & corr != -1 & StimulusType == "WORD"),
  formula = bf(corr ~ Condition * Gender + (Condition * Gender | PROLIFIC_PID) + (Condition | Target)),
  family = bernoulli(), chains = 4, iter = 5000, backend = "cmdstanr"
)
```

</div>

::: {.callout-tip}
Models were run separately via `analysis_exp2.R` and saved with **qs2** for fast loading. No need to re-run MCMC in the presentation!
:::

## Results: RT Model (Ex-Gaussian)

```{r}
#| cache: true
#| label: tbl-rt
parameter_names <- c(
  "Intercept",
  "Intercept",
  "Concordant vs. Discordant Flanker",
  "Grammatical Gender (Feminine vs. Masculine)",
  "Concordant vs. Discordant:Grammatical Gender",
  "Concordant vs. Discordant Flanker",
  "Grammatical Gender (Feminine vs. Masculine)",
  "Concordant vs. Discordant:Grammatical Gender"
)

blmm_exp2_rt_dist_summary <- cbind(parameter_names, summary(blmm_exp2_rt_dist)$fixed %>% select(-Rhat, -ends_with("_ESS")))

colnames(blmm_exp2_rt_dist_summary) <- c("parameter", "estimate", "SE", "lower", "upper")

beta_or_not <- if_else(rownames(blmm_exp2_rt_dist_summary) %>% startsWith("beta_"), true = "beta", "mu")

gt(blmm_exp2_rt_dist_summary, rowname_col = "parameter") %>%
  tab_row_group(id = "mu", label = md("Parameter: $\\mu$"), rows = which(beta_or_not == "mu")) %>%
  tab_row_group(id = "beta", label = md("Parameter: $\\beta$"), rows = which(beta_or_not == "beta")) %>%
  row_group_order(c("mu", "beta")) %>%
  cols_label(
    parameter = "Parameter",
    estimate = "b",
    SE = "SE",
    lower = "2.5%",
    upper = "97.5%"
  ) %>%
  tab_spanner(
    label = "Estimate",
    columns = c(estimate, SE)
  ) %>%
  tab_spanner(
    label = "Credible Interval (b)",
    columns = c(lower, upper)
  ) %>%
  fmt_number(
    columns = c(estimate, SE, lower, upper),
    decimals = 3
  ) %>% tab_style(
    style = list(
      cell_text(weight = "bold")),
    locations = cells_body(
      columns = c(estimate, SE, lower, upper),
      rows = ((lower > 0 & upper > 0) | (lower < 0 & upper < 0))
    )) %>%
  tab_source_note(source_note = md("Note: *b* = coefficient estimate; *SE* = standard error; *2.5%* and *97.5%* = lower and upper bounds of the 95% credible interval. Coefficients for which the 95% credible interval does not include 0 are shown in **bold**."))
```

## Results: Accuracy Model (Bernoulli)

```{r}
#| cache: true
#| label: tbl-acc
parameter_names_acc <- c(
  "Intercept",
  "Concordant vs. Discordant Flanker",
  "Grammatical Gender (Feminine vs. Masculine)",
  "Concordant vs. Discordant:Grammatical Gender"
)

blmm_acc_exp2_summary <- cbind(parameter_names_acc, summary(blmm_acc_exp2)$fixed %>% select(-Rhat, -ends_with("_ESS")))

colnames(blmm_acc_exp2_summary) <- c("parameter", "estimate", "SE", "lower", "upper")

gt(blmm_acc_exp2_summary, rowname_col = "parameter") %>%
  cols_label(
    parameter = "Parameter",
    estimate = "b",
    SE = "SE",
    lower = "2.5%",
    upper = "97.5%"
  ) %>%
  tab_spanner(
    label = "Estimate",
    columns = c(estimate, SE)
  ) %>%
  tab_spanner(
    label = "Credible Interval (b)",
    columns = c(lower, upper)
  ) %>%
  fmt_number(
    columns = c(estimate, SE, lower, upper),
    decimals = 3
  ) %>% tab_style(
    style = list(
      cell_text(weight = "bold")),
    locations = cells_body(
      columns = c(estimate, SE, lower, upper),
      rows = ((lower > 0 & upper > 0) | (lower < 0 & upper < 0))
    )) %>%
  tab_source_note(source_note = md("Note: *b* = log-odds coefficient estimate; *SE* = standard error; *2.5%* and *97.5%* = lower and upper bounds of the 95% credible interval. Coefficients for which the 95% credible interval does not include 0 are shown in **bold**."))
```

## Why Bayesian? {.smaller}

| Feature | Frequentist (lme4) | Bayesian (brms) |
|---------|-------------------|-----------------|
| Output | p-values, confidence intervals | Posterior distributions |
| Interpretation | "If H₀ is true, probability of data ≥ this extreme" | "Probability distribution over parameter values" |
| Uncertainty | Asymptotic confidence intervals | Full credible intervals |
| Complex models | May fail to converge | More robust via MCMC |
| Prior information | Not used | Can incorporate prior knowledge |
| Computation | Fast | Slower (MCMC sampling) |

. . .

Docker makes Bayesian analyses **accessible** — no need to manually install Stan!

## Results: Visualizing the Data

Lines show individual participant means; points and bold lines show group means.

```{r}
#| cache: true
library(ggplot2)
library(tidyverse)

# Read and filter data for visualization
exp2_all_csv_content <- fs::dir_ls(path = "../data", glob = "*.csv") %>%
  map_dfr(read_csv, .id = "source", col_type = cols(
    .default = col_character(), rt = col_double(), corr = col_integer(), TrialID = col_integer()))

exp2_plot_data <- exp2_all_csv_content |>
  group_by(PROLIFIC_PID) |>
  mutate(counterbalancing_condition = counterbalancing_condition[1]) |>
  ungroup() |>
  filter(!is.na(TrialID) & TrialID < 2000) |>
  filter(rt > .25 & rt < 1.8 & corr == 1 & StimulusType == "WORD") |>
  mutate(Condition = factor(Condition, levels = c("AGR", "DIS")),
         Gender = factor(Gender, levels = c("FEM", "MAS")))

# Plot mean RT by Condition and Gender
participant_means <- exp2_plot_data |>
  group_by(PROLIFIC_PID, Condition, Gender) |>
  summarise(mean_rt = mean(rt, na.rm = TRUE), .groups = "drop")

ggplot(participant_means, aes(Condition, mean_rt, color = Gender)) +
  geom_line(
    aes(group = interaction(PROLIFIC_PID, Gender)),
    linewidth = 0.4, alpha = 0.2, position = position_dodge(0.1)
  ) +
  stat_summary(
    aes(group = Gender), fun = mean, geom = "line",
    linewidth = 1.2, position = position_dodge(0.2)
  ) +
  stat_summary(
    aes(group = Gender), fun = mean, geom = "point",
    size = 3, position = position_dodge(0.2)
  ) +
  labs(x = "Condition (AGR = concordant, DIS = discordant)",
       y = "Response times (s)", color = "Gender") +
  theme_minimal(base_size = 14)
```

## Post-hoc Comparisons with emmeans

```{r}
#| cache: true
library(emmeans)

# Extract estimated marginal means from the Bayesian RT model
emm <- emmeans(blmm_exp2_rt_dist, ~ Condition * Gender)

# Pairwise comparisons: concordant vs discordant at each gender
contrast(emm, method = "pairwise", by = "Gender")
```

. . .

Concordant flankers **credibly speed up** response times for feminine and masculine targets.

# Reproducible Manuscripts

## The Open Science Case

::: {.incremental}
- @munafo2017manifesto: need for open data and transparent analysis workflows
- @ince2012case: computational results cannot be evaluated without the code
- @lindsay2023plea: computational reproducibility must be verified *before* publishing
- @morey2016peer: reviewers should not evaluate manuscripts without data and analysis pipeline
:::

## Quarto for Reproducible Manuscripts

**Quarto** (<https://quarto.org>) combines:

:::: {.columns}

::: {.column width="50%"}
**Prose** (Markdown)

The concordant condition
was **X.X** ms faster
than discordant (*p* < .001).
:::

::: {.column width="50%"}
**Code** (R or Python chunks)
```r
#| label: tbl-summary
#| tbl-cap: "Descriptive statistics"

data |>
  group_by(Condition, Gender) |>
  summarise(M = mean(rt), SD = sd(rt)) |>
  flextable()
```
:::

::::

. . .

- Output formats: HTML, PDF (via LaTeX), Word
- Compatible with **apaquarto** [@schneider2025apaquarto] for APA-formatted manuscripts

## Docker + Quarto = Fully Reproducible Manuscripts

::: {.incremental}
1. Write your manuscript as a `.qmd` file with embedded R code
2. Build and push a Docker image with your exact R environment
3. Any reader can pull the image and compile the manuscript, reproducing every number
4. The Rocker **verse** image includes Quarto + TinyTeX out of the box
5. Our "Analysis in a Box" image adds all the packages needed for apaquarto manuscripts
:::

::: {.callout-important}
This is the gold standard for reproducibility: the image is a **permanent record** of the software environment used to produce your results.
:::

## Manuscript Structure Example {.smaller}

```
my_project/
├── docker-compose.yml          # Container configuration
├── manuscript.qmd              # Main Quarto manuscript
├── _introduction.qmd           # Chapter: Introduction
├── _methods.qmd                # Chapter: Methods
├── _analysis.qmd               # Chapter: Analysis (with R code)
├── _results.qmd                # Chapter: Results
├── references.bib              # Bibliography
├── data/
│   └── experiment_data.csv     # Raw data
└── outputs/
    ├── model_rt.qs2            # Cached model results
    └── figures/                # Generated figures
```

::: {.callout-tip}
Cache computationally intensive operations (like brms models) with `cache: true` or save them with `qs2::qs_save()` so the manuscript compiles quickly on subsequent runs.
:::

# Customizing Your Container

## Building on an Existing Image

To add packages to the base image, create a `Dockerfile`:

```dockerfile
FROM bangele1/analysis-in-a-box-rocker:latest

# Install additional R packages
RUN R -e "install.packages(c('mgcv', 'gratia', 'itsadug'))"

# Install additional LaTeX packages
RUN tlmgr install booktabs multirow
```

Build and push your custom image:

```bash
docker build -t myusername/my-analysis-image:latest .
docker push myusername/my-analysis-image:latest
```

## Docker Compose with More Options

```yaml
services:
  analysis-in-a-box:
    image: bangele1/analysis-in-a-box-rocker:latest
    environment:
      - PASSWORD=yourpassword
      - ROOT=true          # Enable sudo inside container
    ports:
      - "8787:8787"
    volumes:
      - .:/home/rstudio    # Working directory
      - ./data:/data:ro    # Read-only data directory
    cpus: "4"              # Limit CPU usage (standard Compose)
    mem_limit: 8g          # Limit memory usage (standard Compose)
```

## Key Docker Commands

| Command | Effect |
|---------|--------|
| `docker compose up -d` | Start container in background |
| `docker compose down` | Stop and remove container |
| `docker compose logs` | View container logs |
| `docker ps` | List running containers |
| `docker images` | List downloaded images |
| `docker pull image:tag` | Download an image |
| `docker build -t name .` | Build image from Dockerfile |
| `docker push name:tag` | Push image to Docker Hub |

# Devcontainers & AI Agents

## Dev Containers in VS Code {.smaller}

<div style="font-size:0.85em;">

**Dev Containers** let you open any project inside a fully configured Docker container directly in VS Code.

- Uses a `.devcontainer/devcontainer.json` file to define the environment
- VS Code connects to the container — full IntelliSense, debugging, terminal
- Same environment on every machine, every collaborator
- Also works with **GitHub Codespaces** — a full VS Code dev environment in the browser, powered by the same container

```json
{
  "name": "R Analysis",
  "image": "bangele1/analysis-in-a-box-rocker:latest",
  "customizations": {
    "vscode": {
      "extensions": [
        "REditorSupport.r",
        "quarto.quarto",
        "ms-toolsai.jupyter"
      ]
    }
  }
}
```

</div>

## Dev Containers + AI Agents {.smaller}

<div style="font-size:0.85em;">

Dev containers unlock powerful **AI-assisted coding workflows** in VS Code:

:::: {.columns}

::: {.column width="50%"}
**GitHub Copilot / Copilot Chat**

- Autocomplete R, Stan, and Quarto code
- Ask questions about your data or model output inline
- Works inside the container — same packages as your analysis
:::

::: {.column width="50%"}
**AI Agent Tools (e.g., Claude Code, opencode)**

- Autonomous agents that read, edit, and run code
- Install packages, fix errors, write tests — all inside the container
- Reproducible: the agent works in your exact environment
:::

::::

::: {.callout-important}
**Security**: AI agents confined to a devcontainer cannot access your host filesystem, credentials, or other projects. The container acts as a sandbox — agents can only see and modify files you mount into it.
:::

::: {.callout-tip}
Combine `devcontainer.json` with a Docker image to give AI agents a **fully specified environment** — no ambiguity about package versions or system dependencies.
:::

</div>

# Beyond R: OpenSesame Experiments

## Testing Experiments with Docker

Docker isn't just for R analyses — use it to **test behavioral experiments** before deploying in the lab.

<div style="font-size:0.8em;">

:::: {.columns}

::: {.column width="70%"}
**OpenSesame Test Container**

- Pre-built image: <https://hub.docker.com/r/bangele1/opensesame-test-container>
- OpenSesame + PsychoPy + PyGaze
- Virtual desktop via NoVNC — access the GUI in your browser
- Runs on **amd64 and arm64** (Apple Silicon)
:::

::: {.column width="30%"}
**Use cases**

- Test experiments before running participants
- Share reproducible experiment environments with collaborators
- Run on any machine — no local install needed
:::

::::

</div>

## OpenSesame: Getting Started {.smaller}

<div style="font-size:0.85em;">

Create a `docker-compose.yml` in your experiment folder:

```yaml
services:
  opensesame-dev:
    image: bangele1/opensesame-test-container
    volumes:
      - ..:/workspaces/experiments
    ports:
      - "8080:8080"
    cap_add:
      - SYS_ADMIN
    restart: unless-stopped
```

Then run:

```bash
docker compose up -d
```

Open **<http://localhost:8080>** in your browser to access the virtual desktop with OpenSesame.

</div>

::: {.callout-tip}
The container uses `Xvfb` + `x11vnc` + `novnc` to render the GUI in the browser — no display server needed on the host machine.
:::

# Summary

## What We've Covered

<div style="font-size:0.8em;">

::: {.incremental}
1. **The Problem**: Complex R environments are hard to install and replicate
2. **Docker**: Containers provide isolated, reproducible environments
3. **Available Images**: Rocker project and b-data images for R + RStudio
4. **Running the Container**: One command to get a full R/RStudio/brms environment
5. **Analyses**: lme4/buildmer for frequentist, brms/CmdStanR for Bayesian models
6. **Caching**: qs2 for fast loading of pre-computed brms models
7. **Reproducible Manuscripts**: Quarto + Docker = fully reproducible research
8. **Customization**: Docker Compose and Dockerfiles for advanced use cases
9. **Dev Containers + AI Agents**: VS Code dev containers with Copilot and AI coding tools
:::

</div>

## Getting Started Today

:::: {.columns}

::: {.column width="50%"}
**Pull the image:**
```bash
docker run \
  -e PASSWORD=yourpassword \
  -p 8787:8787 \
  -v $(pwd):/home/rstudio \
  bangele1/analysis-in-a-box-rocker:latest
```

Then open <http://127.0.0.1:8787>
:::

::: {.column width="50%"}
**Resources:**

- 📦 Docker Hub: [bangele1/analysis-in-a-box-rocker](https://hub.docker.com/r/bangele1/analysis-in-a-box-rocker)
- 💻 GitHub: [ismaelgutier/analysis_with_docker](https://github.com/ismaelgutier/analysis_with_docker)
- 🐳 Docker Desktop: [docs.docker.com](https://docs.docker.com/get-started/introduction/get-docker-desktop/)
- 📖 Rocker Project: [rocker-project.org](https://rocker-project.org)
:::

::::

## References {.smaller}

::: {#refs}
:::
