---
title: "Analysis in a Box"
subtitle: "A Practical Guide to Using RStudio with Docker"
author:
  - name: Ismael Gutiérrez-Cordero
    affiliation: Universidad de Málaga
  - name: Bernhard Angele
    affiliation: Universidad Antonio de Nebrija
date: today
format:
  revealjs:
    theme: default
    slide-number: true
    toc: true
    toc-depth: 1
    toc-title: "Overview"
    logo: ""
    footer: "Analysis in a Box | RStudio + Docker"
    code-fold: false
    code-tools: true
    highlight-style: github
    transition: slide
    incremental: false
    chalkboard: true
bibliography: references.bib
---

# The Problem

## Modern Analyses Are Complex

::: {.incremental}
- Behavioral and neuroscientific data analysis has become **increasingly complex**
- ANOVAs are no longer sufficient — researchers now use:
  - (Generalized) Linear Mixed Models (GLMMs) via **lme4** [@bates2015fitting]
  - Bayesian multilevel models via **brms** [@burkner2017brms]
- *brms* requires a large number of dependencies:
  - Stan [@stan2025] for MCMC sampling
  - CmdStanR [@gabry2025] (recommended) or RStan [@rstan2025]
  - A modern C++ compiler (RTools on Windows, XCode on Mac)
:::

## Installation Challenges

:::: {.columns}

::: {.column width="50%"}
**New Users**

- Overwhelmed by complex installation chains
- Multiple tools to install before writing a single line of R
- Different instructions per operating system
:::

::: {.column width="50%"}
**Experienced Users**

- Significant time spent re-installing and updating packages
- Package updates can break existing code
- Replicating analyses from even a few years ago can fail
:::

::::

## The Reproducibility Gap

> *"Computational results cannot be comprehensively evaluated without knowing the code used to perform the computations."*
> — @ince2012case

::: {.incremental}
- Package versions change constantly, breaking analysis code
- Solutions like **renv** [@renv] require significant effort from both authors and readers
- We need a simpler, more robust approach...
:::

# Docker to the Rescue

## What Is Docker?

::: {.callout-note}
**Docker** is an open-source platform that allows users to run applications in isolated environments called **containers**.
:::

Key concepts:

| Term | Meaning |
|------|---------|
| **Image** | A pre-built software package with all code and configuration |
| **Container** | A running instance of an image — isolated from the host |
| **Docker Hub** | Public repository for sharing Docker images |

## Why Docker for R Analyses?

::: {.incremental}
- **Consistency**: The same image runs identically on any machine
- **Reproducibility**: Freeze your exact software environment as an image
- **Ease of setup**: One command to get a fully working R + RStudio + brms environment
- **Shareability**: Push your image to Docker Hub; anyone can pull and run it
- **Versioning**: Images can build on each other — add packages to an existing base
:::

## Available Base Images

:::: {.columns}

::: {.column width="50%"}
**🐳 Rocker Project** [@rocker2017]

- <https://rocker-project.org>
- R + RStudio + tidyverse/verse
- Versioned by R version
- Best for **Intel/AMD64** (Windows/Linux)
- `rocker/verse:4.5.3`
:::

::: {.column width="50%"}
**🐳 b-data Images**

- <https://gitlab.com/b-data>
- R + RStudio + tidyverse/verse  
- Supports **both amd64 AND arm64**
- Best for **Apple M-series** and ARM Windows
- Recommended for maximum performance on ARM
:::

::::

## Our "Analysis in a Box" Image

We built a custom image on top of Rocker verse that includes:

- base R + RStudio Server
- tidyverse and data processing packages
- **brms** with CmdStanR (pre-compiled!)
- lme4, buildmer, emmeans
- Quarto + apaquarto template
- LaTeX / TinyTeX for PDF rendering

**Available on Docker Hub:**  
<https://hub.docker.com/r/bangele1/analysis-in-a-box-rocker>

# Getting Started

## Prerequisites

1. **Install Docker Desktop**

   Follow the official guide: <https://docs.docker.com/get-started/introduction/get-docker-desktop/>

2. **Verify Docker is working:**

   ```bash
   docker run -d -p 8081:80 docker/welcome-to-docker
   ```

   Then open <http://localhost:8081> in your browser.

3. **Create a working folder** (where your files will be saved):

   ::: {.panel-tabset}
   ### Windows CMD
   ```cmd
   mkdir %USERPROFILE%\Documents\analysis_box
   cd %USERPROFILE%\Documents\analysis_box
   ```

   ### macOS / Linux
   ```bash
   mkdir -p ~/Documents/analysis_box
   cd ~/Documents/analysis_box
   ```
   :::

## Running the Container

Navigate to your working folder and run:

::: {.panel-tabset}

### Windows CMD

```cmd
docker run -e PASSWORD=yourpassword -p 8787:8787 ^
  -v "%cd%":/home/rstudio ^
  bangele1/analysis-in-a-box-rocker:latest
```

### macOS / Linux

```bash
docker run -e PASSWORD=yourpassword -p 8787:8787 \
  -v $(pwd):/home/rstudio \
  bangele1/analysis-in-a-box-rocker:latest
```

:::

. . .

What do these options mean?

- **`-e PASSWORD=yourpassword`** — sets the RStudio Server login password
- **`-p 8787:8787`** — exposes port 8787 to your browser
- **`-v $(pwd):/home/rstudio`** — mounts your local folder inside the container

## Accessing RStudio

1. Open your browser and navigate to **<http://127.0.0.1:8787>**

2. Log in with:
   - **Username:** `rstudio`
   - **Password:** *(whatever you set after `PASSWORD=`)*

3. You now have a **fully working RStudio environment** with brms, CmdStanR, and all dependencies pre-installed!

::: {.callout-tip}
The CLI/terminal window must remain open for the container to stay running.
:::

## Downloading the Example Files

Inside RStudio, open the **Terminal** tab and run:

```bash
git clone https://github.com/ismaelgutier/analysis_with_docker.git
```

This downloads the example data and analysis scripts.  
Open `analysis_with_docker.RProj` to load the project in RStudio.

## Using Docker Compose

For more complex setups, use **Docker Compose** instead of a long `docker run` command.

Create a file called `docker-compose.yml`:

```yaml
services:
  analysis-in-a-box:
    image: bangele1/analysis-in-a-box-rocker:latest
    environment:
      - PASSWORD=yourpassword
    ports:
      - "8787:8787"
    volumes:
      - .:/home/rstudio
```

Then run:

```bash
docker compose up -d     # Start container in background
docker compose down      # Stop and remove container
```

# Analyses with Docker

## Example Dataset

We replicate analyses from @angele2023does — a lexical decision experiment examining masked priming effects:

:::: {.columns}

::: {.column width="50%"}
**Design**

- **Prime-target relationship**: related vs. unrelated
- **Prime duration**: 33 ms, 50 ms, 67 ms
- Participants and items crossed (fully counterbalanced)
:::

::: {.column width="50%"}
**Outcome measures**

- Response times (RT) in ms
- Accuracy (proportion correct)

**Key question:** Do related primes speed up target recognition?
:::

::::

## Mixed-Effects Models: Why?

Standard ANOVA treats participants (or items) as the only random source of variability.  
Mixed models include **random effects** for both:

:::: {.columns}

::: {.column width="33%"}
**Random intercepts only**

Each participant has a different baseline RT, but the *effect* of condition is the same for everyone.

```r
RT ~ condition + 
  (1 | participant) + 
  (1 | item)
```
:::

::: {.column width="33%"}
**Random intercepts + slopes**

Each participant also has a different *effect* of condition.

```r
RT ~ condition + 
  (condition | participant) +
  (condition | item)
```
:::

::: {.column width="33%"}
**Why it matters**

- Avoids pseudoreplication
- By-item random effects: items differ in difficulty
- Converges to correct Type I error rates
:::

::::

## Frequentist Analysis: lme4 + buildmer

```{r}
#| cache: true
#| 
library(lme4)
library(buildmer)

# Load and filter data
data <- readr::read_csv(
  "../data/data_exp1.csv"
) |>
  dplyr::filter(acc != -1)

# Fit maximal random effects model using buildmer
# (automatically finds the largest converging model)
model_rt <- buildmer(
  RT ~ condition * prime_duration +
    (condition * prime_duration | participant) +
    (condition * prime_duration | target),
  data = data,
  buildmerControl = buildmerControl(direction = "order")
)

summary(model_rt)
```

::: {.callout-note}
**buildmer** [@buildmer] automatically reduces the random effects structure to find the maximal converging model — saving considerable trial and error.
:::

## Bayesian Analysis: brms

```{r}
#| cache: true
#| 
library(brms)

prior_exp1 <- c(set_prior("normal(0,1)", class = "b"))

blmm_exp1_rt_dist <- brm(data = exp1_data_to_include %>% filter(corr == 1 & StimulusType == "Word" & rt > 250 & rt < 1800), 
                    bf(rt ~ Condition * PrimeDuration + (1 + Condition * PrimeDuration|source) + (1 + Condition * PrimeDuration|Target), beta ~ Condition * PrimeDuration + (1 + Condition * PrimeDuration|source) + (1 + Condition * PrimeDuration|Target)),
                    warmup = 1000,
                    iter = 5000,
                    chains = 4,
                    prior = prior_exp1,
                    family = exgaussian(),
                    inits = "random",
                    control = list(adapt_delta = 0.8),
                    cores = 4, backend = "cmdstanr", threads = threading(4))                

# Fit Bayesian LMM for RT
model_brms_rt <- brm(
  log(RT) ~ condition * prime_duration +
    (condition * prime_duration | participant) +
    (condition * prime_duration | target),
  data   = data,
  family = gaussian(),
  chains = 4,
  iter   = 4000,
  cores  = 4,
  backend = "cmdstanr",   # Use CmdStanR backend
  threads = threading(2),
  silent = 0
)
```

::: {.callout-tip}
**CmdStanR** is recommended over RStan: it supports the latest Stan features including **hyperthreading** (`threads_per_chain`) for faster sampling.
:::

## Why Bayesian? {.smaller}

| Feature | Frequentist (lme4) | Bayesian (brms) |
|---------|-------------------|-----------------|
| Output | p-values, confidence intervals | Posterior distributions |
| Interpretation | "If H₀ is true, probability of data ≥ this extreme" | "Probability distribution over parameter values" |
| Uncertainty | Asymptotic confidence intervals | Full credible intervals |
| Complex models | May fail to converge | More robust via MCMC |
| Prior information | Not used | Can incorporate prior knowledge |
| Computation | Fast | Slower (MCMC sampling) |

. . .

Docker makes Bayesian analyses **accessible** — no need to manually install Stan!

## Results: Visualising the Data

```r
library(ggplot2)
library(papaja)

# Plot mean RT by condition and participant
participant_rel <- data |>
  dplyr::filter(stimulus_type == "Word") |>
  group_by(participant, condition) |>
  summarise(mean_rt = mean(RT, na.rm = TRUE), .groups = "drop")

ggplot(participant_rel, aes(condition, mean_rt)) +
  geom_line(
    aes(group = participant, color = as.factor(participant)),
    linewidth = 0.6, alpha = 0.4
  ) +
  stat_summary(
    aes(group = 1), fun = mean, geom = "line",
    linewidth = 1, color = "black"
  ) +
  stat_summary(
    aes(group = 1), fun = mean, geom = "point",
    size = 2, color = "black"
  ) +
  labs(x = "Prime-target relationship", y = "Response times (ms)") +
  papaja::theme_apa() +
  theme(legend.position = "none")
```

## Post-hoc Comparisons with emmeans

```r
library(emmeans)

# Extract estimated marginal means
emm <- emmeans(model_rt, ~ condition * prime_duration)

# Pairwise comparisons: related vs. unrelated at each prime duration
contrast(emm, method = "pairwise", by = "prime_duration")
```

Output:
```
prime_duration = 33ms:
  contrast              estimate   SE  df t.ratio p.value
  related - unrelated     -18.3  4.2 189  -4.36   <.001

prime_duration = 50ms:
  contrast              estimate   SE  df t.ratio p.value
  related - unrelated     -22.1  4.5 191  -4.91   <.001
```

. . .

Related primes **significantly reduced** response times compared to unrelated primes.

# Reproducible Manuscripts

## The Open Science Case

::: {.incremental}
- @munafo2017manifesto: need for open data and transparent analysis workflows
- @ince2012case: computational results cannot be evaluated without the code
- @lindsay2023plea: computational reproducibility must be verified *before* publishing
- @morey2016peer: reviewers should not evaluate manuscripts without data and analysis pipeline
:::

## Quarto for Reproducible Manuscripts

**Quarto** (<https://quarto.org>) combines:

:::: {.columns}

::: {.column width="50%"}
**Prose** (Markdown)
```markdown
The related prime condition 
was `r round(estimate, 1)` ms 
faster than unrelated 
(*p* < .001).
```
:::

::: {.column width="50%"}
**Code** (R or Python chunks)
```r
#| label: tbl-summary
#| tbl-cap: "Descriptive statistics"

data |>
  group_by(condition) |>
  summarise(M = mean(RT), SD = sd(RT)) |>
  flextable()
```
:::

::::

. . .

- Output formats: HTML, PDF (via LaTeX), Word
- Compatible with **apaquarto** [@schneider2025apaquarto] for APA-formatted manuscripts

## Docker + Quarto = Fully Reproducible Manuscripts

::: {.incremental}
1. Write your manuscript as a `.qmd` file with embedded R code
2. Build and push a Docker image with your exact R environment
3. Any reader can pull the image and compile the manuscript, reproducing every number
4. The Rocker **verse** image includes Quarto + TinyTeX out of the box
5. Our "Analysis in a Box" image adds all the packages needed for apaquarto manuscripts
:::

::: {.callout-important}
This is the gold standard for reproducibility: the image is a **permanent record** of the software environment used to produce your results.
:::

## Manuscript Structure Example {.smaller}

```
my_project/
├── docker-compose.yml          # Container configuration
├── manuscript.qmd              # Main Quarto manuscript
├── _introduction.qmd           # Chapter: Introduction
├── _methods.qmd                # Chapter: Methods
├── _analysis.qmd               # Chapter: Analysis (with R code)
├── _results.qmd                # Chapter: Results
├── references.bib              # Bibliography
├── data/
│   └── experiment_data.csv     # Raw data
└── outputs/
    ├── model_rt.qs             # Cached model results
    └── figures/                # Generated figures
```

::: {.callout-tip}
Cache computationally intensive operations (like brms models) with `cache: true` or save them with `qs2::qsave()` so the manuscript compiles quickly on subsequent runs.
:::

# Customizing Your Container

## Building on an Existing Image

To add packages to the base image, create a `Dockerfile`:

```dockerfile
FROM bangele1/analysis-in-a-box-rocker:latest

# Install additional R packages
RUN R -e "install.packages(c('mgcv', 'gratia', 'itsadug'))"

# Install additional LaTeX packages
RUN tlmgr install booktabs multirow
```

Build and push your custom image:

```bash
docker build -t myusername/my-analysis-image:latest .
docker push myusername/my-analysis-image:latest
```

## Docker Compose with More Options

```yaml
services:
  analysis-in-a-box:
    image: bangele1/analysis-in-a-box-rocker:latest
    environment:
      - PASSWORD=yourpassword
      - ROOT=true          # Enable sudo inside container
    ports:
      - "8787:8787"
    volumes:
      - .:/home/rstudio    # Working directory
      - ./data:/data:ro    # Read-only data directory
    cpus: "4"              # Limit CPU usage (standard Compose)
    mem_limit: 8g          # Limit memory usage (standard Compose)
```

## Key Docker Commands

| Command | Effect |
|---------|--------|
| `docker compose up -d` | Start container in background |
| `docker compose down` | Stop and remove container |
| `docker compose logs` | View container logs |
| `docker ps` | List running containers |
| `docker images` | List downloaded images |
| `docker pull image:tag` | Download an image |
| `docker build -t name .` | Build image from Dockerfile |
| `docker push name:tag` | Push image to Docker Hub |

# Summary

## What We've Covered

::: {.incremental}
1. **The Problem**: Complex R environments are hard to install and replicate
2. **Docker**: Containers provide isolated, reproducible environments
3. **Available Images**: Rocker project and b-data images for R + RStudio
4. **Running the Container**: One command to get a full R/RStudio/brms environment
5. **Analyses**: lme4/buildmer for frequentist, brms/CmdStanR for Bayesian models
6. **Reproducible Manuscripts**: Quarto + Docker = fully reproducible research
7. **Customization**: Docker Compose and Dockerfiles for advanced use cases
:::

## Getting Started Today

:::: {.columns}

::: {.column width="50%"}
**Pull the image:**
```bash
docker run \
  -e PASSWORD=yourpassword \
  -p 8787:8787 \
  -v $(pwd):/home/rstudio \
  bangele1/analysis-in-a-box-rocker:latest
```

Then open <http://127.0.0.1:8787>
:::

::: {.column width="50%"}
**Resources:**

- 📦 Docker Hub: [bangele1/analysis-in-a-box-rocker](https://hub.docker.com/r/bangele1/analysis-in-a-box-rocker)
- 💻 GitHub: [ismaelgutier/analysis_with_docker](https://github.com/ismaelgutier/analysis_with_docker)
- 🐳 Docker Desktop: [docs.docker.com](https://docs.docker.com/get-started/introduction/get-docker-desktop/)
- 📖 Rocker Project: [rocker-project.org](https://rocker-project.org)
:::

::::

## References {.smaller}

::: {#refs}
:::
